# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r}
datRaw <- read.table('activity.csv', header=TRUE, sep=',', colClasses=c('numeric', 'Date', 'integer'))

```


## What is mean total number of steps taken per day?

```{r, fig.cap='Histogram of number of steps per day'}
library(ggplot2)
myTbyDay <- aggregate(steps~date, data=datRaw, FUN='sum', na.action='na.omit')
ggplot(myTbyDay, aes(x=steps)) +
    geom_histogram(binwidth=1000)
```

```{r}
mean(myTbyDay$steps)
median(myTbyDay$steps)
```

The mean number of steps per day is `r format(mean(myTbyDay$steps), nsmall=2)`.
The median number of steps per day is `r format(median(myTbyDay$steps), nsmall=2)`.

## What is the average daily activity pattern?

```{r, fig.cap='Time series of average number of steps by 5 minutres interval'}
myTbyMin <- aggregate(steps~interval, data=datRaw, FUN='mean', na.action='na.omit')

ggplot(myTbyMin, aes(x=interval, y=steps)) +
    geom_line()
myTbyMin$interval[which.max(myTbyMin$steps)]
```

The 5 minute interval with maximum average number of steps is interval `r myTbyMin$interval[which.max(myTbyMin$steps)]`.

## Imputing missing values

```{r}
writeLines('Simple :')
sum(is.na(datRaw$steps))
writeLines('More subtle :')
sum(!complete.cases(datRaw))
```

There are `r sum(is.na(datRaw$steps))` missing values in the dataset.

We fill missing values with the mean for the 5 minute interval :

```{r}
datComplete <- datRaw
indexOfMissingData <- is.na(datRaw$steps)
datComplete$steps[indexOfMissingData] <- rep(myTbyMin$steps, 61)[indexOfMissingData]
```

Histogram and mean / median of the data set with missing values filled in :

```{r, fig.cap='Histogram with missing data filled in'}
myTbyDayComplete <- aggregate(steps~date, data=datComplete, FUN='sum', na.action='na.omit')
ggplot(myTbyDayComplete, aes(x=steps)) +
    geom_histogram(binwidth=1000)

mean(myTbyDayComplete$steps)
median(myTbyDayComplete$steps)
```

The effect of filling in the missing values has no effect on the global mean, but has an impact on the variance of data (reduced variance, with a more pronounced central pic on the histogram).
The median is now equal to the mean.


## Are there differences in activity patterns between weekdays and weekends?

```{r, fig.cap='Activity pattern for weekdays and weekends'}
datComplete <- within(datComplete, {
                      dayType <- (weekdays(date) %in% c('samedi', 'dimanche')) +1
                      dayType <- factor(dayType, labels=c('weekday', 'weekend'))
})

myTbyMinAndDayType <- aggregate(steps~interval+dayType, data=datComplete, FUN='mean')

ggplot(myTbyMinAndDayType, aes(x=interval, y=steps)) +
    geom_line() +
    facet_grid(dayType ~ .)
```
